---
title: "Philippine Rice Industry  Outlook"
subtitle: "Data stories describing the current situation of the Philippine Rice Industry"
# format: revealjs
format:
  revealjs:
    css: custom-styles.css
    touch: true
    controls: true
    slide-number: true
    transition: fade
    chalkboard: 
      buttons: false
    logo: www/philricelogo2.png
    footer: "PhilRice Data Analytics Center"
execute: 
  echo: false
  warning: false
  message: false
  freeze: auto
---

<!-- ## Important contents -->

```{r}
#| label: data and lib
#| echo: false
#|
library(tidyverse)
library(DT)
library(htmltools)
library(plotly)
library(shiny)
library(htmlwidgets)
library(geojson)
library(geojsonio)
library(geojsonsf)
library(sf)
library(leaflet)


clean_location <-  function(variable){
  case_match(
    str_to_lower(str_squish(variable)),
    c("region i","ilocos","region 1","ilocos region", "region i (ilocos region)","region 1 (ilocos region)") ~ "Region I (Ilocos Region)",
    c("region ii","region 2","cagayan valley", "region ii (cagayan valley)","region 2 (cagayan valley)") ~ "Region II (Cagayan Valley)",
    c("region iii","region 3","central luzon", "region iii (central luzon)","region 3 (central luzon)") ~ "Region III (Central Luzon)",
    c("region iv-a","southern tagalog","region 4a","calabarzon", "region iv-a (calabarzon)","region 4-a (calabarzon)") ~ "Region IV-A (CALABARZON)",
    c("region v","bicol","region 5","bicol region", "region v (bicol region)","region 5 (bicol region)" ) ~ "Region V (Bicol Region)",
    c("region vi","region 6","western visayas", "region vi (western visayas)", "region 6 (western visayas)") ~ "Region VI (Western Visayas)",
    c("region vii","region 7","central visayas", "region vii (central visayas)", "region 7 (central visayas)") ~ "Region VII (Central Visayas)",
    c("region viii","region 8","eastern visayas", "region viii (eastern visayas)","region 8 (eastern visayas)") ~ "Region VIII (Eastern Visayas)",
    c("region ix","region 9","zamboanga peninsula", "region ix (zamboanga peninsula)", "region 9 (zamboanga peninsula)") ~ "Region IX (Zamboanga Peninsula)",
    c("region x","region 10","northern mindanao", "region x (northern mindanao)","region 10 (northern mindanao)") ~ "Region X (Northern Mindanao)",
    c("region xi","region 11","davao region", "region xi (davao region)","region 11 (davao region)") ~ "Region XI (Davao Region)",
    c("region xii","region 12","soccsksargen", "region xii (soccsksargen)","region 12 (soccsksargen)") ~ "Region XII (SOCCSKSARGEN)",
    c("region xiii","region 13","caraga", "region xiii (caraga)","region 13 (caraga region)","REGION XIII (Caraga) ") ~ "Region XIII (Caraga)",
    c("region iv-b","mimaropa", "region iv-b (mimaropa)","mimaropa region","region 4-b (mimaropa region)" ) ~ "MIMAROPA Region",
    c("cordillera administrative region (car)", "car","cordillera administrative region") ~ "Cordillera Administrative Region (CAR)",
    c("autonomous region in muslim mindanao (armm)","armm","barmm","bangsamoro autonomous region in muslim mindanao (barmm)", "bangsamoro autonomous region in muslim mindanao") ~ "Bangsamoro Autonomous Region in Muslim Mindanao (BARMM)",
    c("city of davao","davao city")~"davao city",
    c("cotabato (north cotabato)","north cotabato")~"north cotabato",
    c("compostela valley","davao de oro","davao de oro (compostela valley)")~"davao de oro (compostela valley)",
    c("western samar","samar","samar (western samar")~"samar (western samar)",
    c("maguindanao (excluding cotabato city)") ~ "maguindanao",
    c("basilan (excluding city of isabela)")~"basilan",
    c("ncr")~"national capital region (ncr)",
    c("agusan norte")~"agusan del norte",
    c("agusan sur")~"agusan del sur",
    c("davao norte")~"davao del norte",
    c("davao sur")~"davao del sur",
    c("lanao norte")~"lanao del norte",
    c("lanao sur")~"lanao del sur",
    c("mt. province")~"mountain province",
    c("mindoro occidental") ~"occidental mindoro",
    c("mindoro oriental") ~"oriental mindoro",
    c("surigao norte")~"surigao del norte",
    c("surigao sur")~"surigao del sur",
    c("zamboanga norte")~"zamboanga del norte",
    c("zamboanga sur")~"zamboanga del sur",
    .default = str_to_lower(variable)
  )
}


# sua_ensemb<- readRDS("ensemb_phl_dt2030.rds")


# sua_target <- readRDS("sua_new5_dt_phl.rds")

seedclass_area <-  readRDS("AreabySeedclass.RDS")

area_cropestab <- readRDS("areacropestab.rds")

nis_total <- read_csv("nis_total.csv") |>
  pivot_longer(Dry1_Rice:Wet_tot,
               names_to = c("seas","comm"),
               names_sep ="_",
               values_to = "val")


# nis_df <-  readRDS("nis_df.rds")

province_map <-  readRDS("provincemap.RDS")

```

<!-- ::: {.incremental .tiny-text} -->

<!-- -   Updated Rice SUA and projections (2025-2030) -->

<!--     -   *Source: SUA Dashboard* -->

<!-- -   Rice Areas planted to inbred rice -->

<!--     -   *Source: RBFHS 2021-2022/SCR PUF* 2022 -->

<!--     -   [Ricelytics Seeds and Varieties Dashboard](https://ricelytics.philrice.gov.ph/seeds_and_varieties/) -->

<!-- -   Rice Areas planted to Hybrid rice -->

<!--     -   *Source: RBFHS 2021-2022/SCR PUF* 2022 -->

<!-- -   Rice Areas practicing Direct Seeding -->

<!--     -   *Source: RBFHS 2021-2022/SCR PUF* 2022 (in terms of man days) -->

<!-- -   Rice Areas practicing double dry cropping -->

<!--     -   Source: NIA data program area for double dry 2025 / NIS explorer map -->

<!-- -   Outlook -->

<!-- ::: -->

# Historical trends

## Production and Yield trends in the Philippines {.tiny-text}

```{r}
#| label: ann_prod
#| echo: false
#| warning: false
#| message: false

psadata <-  readRDS("psadata.rds")


nat_annual<- 
  psadata |> 
  filter(PERIOD_CODE==3) |> 
  filter(eco_code==1) |> 
  filter(LOCATION_TYPE=="NATIONAL") |> 
  select(LOCATION,eco_code,date,PRODUCTION:YIELD)



prod_plot <- 
plot_ly(nat_annual,
        x = ~year(date), 
        y = ~PRODUCTION,
        marker = list(color = "darkgreen"),
        name = "Total production (mt)",
        type = 'bar',
        width =1000,
        height=600) |> 
  layout(
         xaxis = list(title = "Year",
                      rangeslider = list(visible = TRUE)),
         yaxis = list(title = "Production in MT"),
         barmode = "overlay")



prod_plot |> 
    add_lines(y=~YIELD, x=~year(date),
            name = "Average yield (mt/ha)",
            data = nat_annual,
            yaxis = 'y2',
            line = list(color = "darkorange", width = 4),
            marker = list(color= "orange")
            ) |> 
    layout(
    title =  "National Annual Production (1971 - Present)",
    yaxis = list(title = "Production (MT)"),
    yaxis2 = list(
      title = "Average yield (mt/ha)",
      overlaying = "y",
      side = "right",
      titlefont = list(size = 12),
      tickfont = list(size = 10),
      standoff = 20  
    ),
    hovermode = 'x unified',
    xaxis = list(title = "Year"),
    legend = list(x = 0.05, y = 0.95),
    margin = list(r = 80)
    ) |> 
    config(displayModeBar=FALSE)
  
```

## Trends in sources of production growth/decline {.tiny-text}

```{r}
#| label: sopg
#| echo: false


sopg <-  read_csv("national_sopg.csv") |> 
  mutate(across(c(ps_area,ps_yield),
                ~(.*100))) |> 
  mutate(ps_area2 = case_when(
    prod_change_total>0 & ps_area<0 & ps_yield>0~0,
    prod_change_total>0 & ps_yield<0 & ps_area>0 ~1,
    prod_change_total<0 & ps_area>0 & ps_yield<0~0,
    prod_change_total<0 & ps_yield>0 & ps_area<0 ~1,
    .default = (ps_area/(ps_yield+ps_area))),
    ps_yield2 = case_when(
      prod_change_total>0 & ps_yield<0 & ps_area>0 ~0,
      prod_change_total>0 & ps_area<0 & ps_yield>0 ~1,
      prod_change_total<0 & ps_yield>0 & ps_area<0~0,
      prod_change_total<0 & ps_area>0 & ps_yield<0 ~1,
      .default = (ps_yield/(ps_yield+ps_area))
    ),
    prod_area_x = prod_change_total*ps_area2,
    prod_yield_x=prod_change_total*ps_yield2,
    area_perc = ps_area2*100,
    yield_perc = ps_yield2*100
  )


plot_ly(x=~Year,
        data = sopg |> filter(Year>2000),
        width=1000,
        height = 600) |> 
  add_bars(
    y = ~prod_area_x,
    marker = list(color = "darkorange"),
    name = "Production change due to area",
    type = 'bar',
    text = ~paste("% due to area:",round(area_perc), "%"),
    textposition="none",
    hoverinfo = 'text') |> 
  add_bars(
    y = ~prod_yield_x,
    marker = list(color = "brown"),
    name = "Production change due to yield",
    type = 'bar',
    text = ~paste("% due to yield:",round(yield_perc), "%"),
    textposition="none",
    hoverinfo = 'text') |>  
    add_lines(
    y = ~prod_change_total,
    line = list(color = "darkgreen"),
    marker = list(color = "darkgreen"),
    text = ~paste("Production Change:",format(round(prod_change_total),big.mark=","), "MT"),
    hoverinfo = 'text',
    name = "Year-on-Year Change in Production") |>
  layout(
    
    xaxis = list(title = "Year",
                 rangeslider = list(visible = TRUE)),
    yaxis = list(title = "Volume in MT"),
    # hovermode = "x unified",
    barmode = "stack",
    legend = list(x = 0.05, y = 1,
                  bgcolor = 'rgba(255, 255, 255, 0.4)'),
    hovermode = 'x unified',
    xaxis = list(title = "Year")
  )|> 
    config(displayModeBar=FALSE)
  
    
  
  

```

# 2024 Performance

## Annual Production {.tiny-text}

::: columns
::: {.column width="30%"}
```{r}
#|label: prod_overall

prod_natl <- 
psadata |> 
  filter(YEAR==2024) |> 
  filter(LOCATION_TYPE=="NATIONAL") |> 
  filter(PERIOD=="ANNUAL") |> 
  mutate(PRODUCTION = round(PRODUCTION/1000000,digits = 2))

prod_natl_alleco<- prod_natl |> 
  filter(eco_code==1) |> 
  pull(PRODUCTION)

prod_natl_irr<- prod_natl |> 
  filter(eco_code==2) |> 
  pull(PRODUCTION)

prod_natl_rain<- prod_natl |> 
  filter(eco_code==3) |> 
  pull(PRODUCTION)



```

::: flex-row
::: {.value-box .blue}
::: value-title
All Ecosystem
:::

::: value-number
`r prod_natl_alleco`
:::

::: value-subtitle
Million MT
:::
:::

::: {.value-box .green}
::: value-title
Irrigated
:::

::: value-number
`r prod_natl_irr`
:::

::: value-subtitle
Million MT
:::
:::

::: {.value-box .orange}
::: value-title
Rainfed
:::

::: value-number
`r prod_natl_rain`
:::

::: value-subtitle
Million MT
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| label: prodmap2
#| echo: false

psa_prov2024 <- 
psadata |> 
  filter(YEAR==2024) |> 
  filter(LOCATION_TYPE=="PROVINCE") |> 
  filter(PERIOD=="ANNUAL") |> 
  filter(eco_code==1) |> 
  mutate(location2 = if_else(LOCATION == "COTABATO",
                             "north cotabato",
                             clean_location(LOCATION))
  )

psa_prodmap <- 
province_map |>
  left_join(psa_prov2024,
            by=c("location"="location2"))



prod_pal <- colorBin(palette =c("#FF7F00",
                                "#FFD92F",
                                "#4DAF4A",
                                "#1F78B4"),
                        na.color = "white",
                        domain = psa_prodmap$PRODUCTION[psa_prodmap$PRODUCTION>0], 
                        bins = c(0,51000,110000,310000,Inf))

psa_prodmap <- 
  psa_prodmap |> 
  mutate(color = if_else(PRODUCTION==0, "grey",
                         prod_pal(PRODUCTION)))

labels_prod <- sprintf("Province: %s<br>Production: %0.2f", 
                          str_to_title(psa_prodmap$location),
                          psa_prodmap$PRODUCTION) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(psa_prodmap,
        width = 700,
        height = 600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor =~color,
    fillOpacity = 1,
    weight = 1,
    opacity = 1,
    color = "black",
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_prod
  ) %>%
    addLegend(pal = prod_pal, 
              values = ~PRODUCTION, 
              title = htmltools::HTML("Annual Palay Production in all ecosystems <br>by province"), 
              
              opacity = 0.7,
              labFormat = function(type, cuts, p) {
                c("<51,000", 
                  "51,000 - 110,000", 
                  "110,000 - 310,000", 
                  "> 310.000")  # matches the bins
              },
              na.label = "No production") |> 
    addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>PSA</div>",
    position = "bottomleft"
  )


```
:::
:::

## Average Yield {.tiny-text}

::: columns
::: {.column width="30%"}
```{r}
#|label: yld_overall

yld_natl <- 
psadata |> 
  filter(YEAR==2024) |> 
  filter(LOCATION_TYPE=="NATIONAL") |> 
  filter(PERIOD=="ANNUAL") |> 
  mutate(YIELD = round(YIELD,digits = 2))

yld_natl_alleco<- yld_natl |> 
  filter(eco_code==1) |> 
  pull(YIELD)

yld_natl_irr<- yld_natl |> 
  filter(eco_code==2) |> 
  pull(YIELD)

yld_natl_rain<- yld_natl |> 
  filter(eco_code==3) |> 
  pull(YIELD)



```

::: flex-row
::: {.value-box .blue}
::: value-title
All Ecosystem
:::

::: value-number
`r yld_natl_alleco`
:::

::: value-subtitle
metric tons per hectare
:::
:::

::: {.value-box .green}
::: value-title
Irrigated
:::

::: value-number
`r yld_natl_irr`
:::

::: value-subtitle
metric tons per hectare
:::
:::

::: {.value-box .orange}
::: value-title
Rainfed
:::

::: value-number
`r yld_natl_rain`
:::

::: value-subtitle
metric tons per hectare
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| echo: false


yield_pal <- colorBin(palette =c("#F1A63C",
                                "#FFF883",
                                "#29883E",
                                "#3B93E4",
                                "#B07AA1"),
                        na.color = "white",
                        domain = psa_prodmap$YIELD[psa_prodmap$YIELD>0], 
                        bins = c(0,3,4,5,6,Inf))

psa_yldmap <- 
  psa_prodmap |> 
  mutate(color = if_else(YIELD==0, "grey",
                         yield_pal(YIELD)))

labels_yld <- sprintf("Province: %s<br>Average Yield: %0.2f", 
                          str_to_title(psa_yldmap$YIELD),
                          psa_yldmap$YIELD) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(psa_yldmap,
        width = 700,
        height = 600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor =~color,
    fillOpacity = 1,
    weight = 1,
    opacity = 1,
    color = "black",
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_yld
  ) %>%
    addLegend(pal = yield_pal, 
              values = ~YIELD, 
              title = htmltools::HTML("Annual Average Yield of Palay<br> in all ecosystems, by province"), 
              
              opacity = 0.7,
              labFormat = function(type, cuts, p) {
                c("<=3", 
                  ">3 to 4", 
                  ">4 to 5", 
                  ">5 to 6",
                  ">6")  # matches the bins
              },
              na.label = "No production") |> 
    addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>PSA</div>",
    position = "bottomleft"
  )


```
:::
:::

# Seed Class Use and Crop Establishment

## Seed Class Use (RBFHS 2022 DS) {.tiny-text}

```{r}

sclass_natl_ds <- 
seedclass_area |> 
  filter(location=="0-philippines") |> 
  filter(data_season=="DS")


sclass_nat_ds_hyb <- 
  sclass_natl_ds |> 
  pull(perce_hybrid_area) |> 
  round(digits = 2)
  
sclass_nat_ds_hqs <- 
  sclass_natl_ds |> 
  pull(perce_highInbred_area) |> 
  round(digits = 2)
  
sclass_nat_ds_lqs<- 
  sclass_natl_ds |> 
  pull(perce_lowInbred_area) |> 
  round(digits = 2)


sclass_natl_ws <- 
seedclass_area |> 
  filter(location=="0-philippines") |> 
  filter(data_season=="WS")


sclass_nat_ws_hyb <- 
  sclass_natl_ws |> 
  pull(perce_hybrid_area) |> 
  round(digits = 2)
  
sclass_nat_ws_hqs <- 
  sclass_natl_ws |> 
  pull(perce_highInbred_area) |> 
  round(digits = 2)
  
sclass_nat_ws_lqs<- 
  sclass_natl_ws |> 
  pull(perce_lowInbred_area) |> 
  round(digits = 2)
    

```

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-title
Hybrid Seeds
:::

::: value-number
`r sclass_nat_ds_hyb` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .green}
::: value-title
High Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ds_hqs` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .orange}
::: value-title
Low Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ds_lqs` %
:::

::: value-subtitle
of total area planted
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| label: hqs_ds
#| echo: false


seedclass_map <- 
province_map |> left_join(seedclass_area |> 
                            filter(year==2022,
                                   locType==2) |> 
                            mutate(location = clean_location(location)) |> 
                            select(location,totarea:perce_lowInbred_area)
                          )


bin_sclass <- quantile(seedclass_map$perce_highInbred_area, probs = seq(0, 1, length.out = 4), na.rm = TRUE, names = FALSE)


pal_sclass <- colorBin(palette = c("#FFA87D",
                                   "#E77561",
                                   "#BC8B77",
                                   "#BC5F3B"
                                   ), 
                       na.color = "white",
                       domain = seedclass_map$perce_highInbred_area, 
                       bins = c(0,25,50,75,100))

seedclass_map <- 
  seedclass_map |> 
  mutate(color = if_else(perce_highInbred_area==0, "white",
                         pal_sclass(perce_highInbred_area)))


labels_sclass <- sprintf("Province: %s<br>Percent: %0.2f", str_to_title(seedclass_map$location), seedclass_map$perce_highInbred_area) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(seedclass_map,
        width = 700,
        height=600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor = ~pal_sclass(perce_highInbred_area),
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_sclass
  ) %>%
    addLegend(pal = pal_sclass, 
              values = ~perce_highInbred_area, 
            title = htmltools::HTML("Percent rice area planted with <br>High Quality Inbred Seeds (2022 DS)"), 
            opacity = 0.7,
              labFormat = function(type, cuts, p) {
    c("<25", "25-50", "50-75", "More than 75")  # matches the bins
  },
  na.label = "No adoption") |> 
    addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022 Survey</div>",
    position = "bottomleft"
  )

```
:::
:::

## Seed Class Use (RBFHS 2022 DS) {.tiny-text}

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-title
Hybrid Seeds
:::

::: value-number
`r sclass_nat_ds_hyb` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .green}
::: value-title
High Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ds_hqs` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .orange}
::: value-title
Low Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ds_lqs` %
:::

::: value-subtitle
of total area planted
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| echo: false


bin_sclass2 <- quantile(seedclass_map$perce_hybrid_area, probs = seq(0, 1, length.out = 4), na.rm = TRUE, names = FALSE)


pal_sclass2 <- colorBin(palette =c("#00F4FF",
                                   "#00C5E9",
                                   "#004CFF",
                                   "#56A3F2"),
                        na.color = "white",
                        domain = seedclass_map$perce_hybrid_area[seedclass_map$perce_hybrid_area>0], 
                        bins = c(0,25,50,75,100))

seedclass_map2 <- 
  seedclass_map |> 
  mutate(color = if_else(perce_hybrid_area==0, "white",
                         pal_sclass2(perce_hybrid_area)))

labels_sclass2 <- sprintf("Province: %s<br>Percent: %0.2f", 
                          str_to_title(seedclass_map2$location),
                          seedclass_map2$perce_hybrid_area) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(seedclass_map2,
        width=700,
        height=600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor =~color,
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_sclass2
  ) %>%
    addLegend(pal = pal_sclass2, 
              values = ~perce_hybrid_area, 
              title = htmltools::HTML("Percent rice area planted with<br> Hybrid Seeds (2022 DS)"), 
              opacity = 0.7,
              labFormat = function(type, cuts, p) {
                c("<25", "25-50", "50-75", "More than 75")  # matches the bins
              },
              na.label = "No adoption") |> 
    addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022 Survey</div>",
    position = "bottomleft"
  )

```
:::
:::

## Seed Class Use (RBFHS 2021 WS) {.tiny-text}

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-title
Hybrid Seeds
:::

::: value-number
`r sclass_nat_ws_hyb` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .green}
::: value-title
High Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ws_hqs` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .orange}
::: value-title
Low Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ws_lqs` %
:::

::: value-subtitle
of total area planted
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| label: hqs_ws
#| echo: false

seedclass_map_ws <- 
province_map |> left_join(seedclass_area |> 
                            filter(year==2021,
                                   locType==2) |> 
                            mutate(location = clean_location(location)) |> 
                            select(location,totarea:perce_lowInbred_area)
                          )


pal_sclass_ws <- colorBin(palette = c("#FFA87D",
                                      "#E77561",
                                      "#BC8B77",
                                      "#BC5F3B"), 
                          na.color = "white",
                          domain = seedclass_map_ws$perce_highInbred_area, 
                          bins = c(0,25,50,75,100))





seedclass_map_ws <- 
  seedclass_map_ws |> 
  mutate(color = if_else(perce_highInbred_area==0, "white",
                         pal_sclass_ws(perce_highInbred_area)))


labels_sclass_ws <- sprintf("Province: %s<br>Percent: %0.2f", str_to_title(seedclass_map_ws$location), seedclass_map_ws$perce_highInbred_area) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(seedclass_map_ws,
        width=700,
        height=600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =6)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor = ~pal_sclass_ws(perce_highInbred_area),
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_sclass
  ) %>%
    addLegend(pal = pal_sclass_ws, 
              values = ~perce_highInbred_area, 
            title = htmltools::HTML("Percent rice area planted with<br>High Quality Inbred Seeds (2021 WS)"), 
            opacity = 0.7,
              labFormat = function(type, cuts, p) {
    c("<25", "25-50", "50-75", "More than 75")  # matches the bins
  },
  na.label = "No adoption") |> 
      addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022</div>",
    position = "bottomleft"
  )
```
:::
:::

## Seed Class Use (RBFHS 2021 WS) {.tiny-text}

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-title
Hybrid Seeds
:::

::: value-number
`r sclass_nat_ws_hyb` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .green}
::: value-title
High Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ws_hqs` %
:::

::: value-subtitle
of total area planted
:::
:::

::: {.value-box .orange}
::: value-title
Low Quality Inbred Seeds
:::

::: value-number
`r sclass_nat_ws_lqs` %
:::

::: value-subtitle
of total area planted
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#| echo: false



pal_sclass2WS <- colorBin(palette =c("#00F4FF",
                                   "#00C5E9",
                                   "#004CFF",
                                   "#56A3F2"),
                        na.color = "white",
                        domain = seedclass_map_ws$perce_hybrid_area[seedclass_map_ws$perce_hybrid_area>0], 
                        bins = c(0,25,50,75,100))

seedclass_map2_WS <- 
  seedclass_map_ws |> 
  mutate(color = if_else(perce_hybrid_area==0, "white",
                         pal_sclass2WS(perce_hybrid_area)))

labels_sclass2WS <- sprintf("Province: %s<br>Percent: %0.2f", 
                          str_to_title(seedclass_map2_WS$location),
                          seedclass_map2_WS$perce_hybrid_area) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(seedclass_map2_WS,
        width=700,
        height = 600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =6)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor =~color,
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels_sclass2WS
  ) %>%
    addLegend(pal = pal_sclass2WS, 
              values = ~perce_hybrid_area, 
              title = htmltools::HTML("Percent rice area planted with<br>Hybrid Seeds (2021 WS)"), 
              opacity = 0.7,
              labFormat = function(type, cuts, p) {
                c("<25", "25-50", "50-75", "More than 75")  # matches the bins
              },
              na.label = "No adoption") |> 
      addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022</div>",
    position = "bottomleft"
  )
```
:::
:::

## Area % planted to Direct Seeding (RBFHS 2022 DS) {.tiny-text}

```{r}
#|echo: false

are_dseed <- 
area_cropestab |>
  filter(location=="0-philippines")

area_dseed_DS <-   
are_dseed |>
  filter(data_season=="DS") |> 
  pull(perce_direct) |> 
  round(digits=2)

area_dssed_WS <- 
are_dseed |>
  filter(data_season=="WS") |> 
  pull(perce_direct) |> 
  round(digits=2)
  


```

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-number
`r area_dseed_DS` %
:::

::: value-title
Area planted with Direct Seeding
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#|echo: false
#|fig-height: 12

cropestab_map <- 
province_map |> left_join(area_cropestab |> 
                            filter(year==2022,
                                   locType==2) |> 
                            mutate(location = clean_location(location)) |> 
                            select(location,totalarea:perce_direct)
                          )


bins <- quantile(cropestab_map$direct, probs = seq(0, 1, length.out = 5), na.rm = TRUE, names = FALSE)

pal <- colorBin(palette =c("#00F4FF",
                            "#00C5E9",
                            "#004CFF",
                            "#56A3F2"
                            ),
                 na.color = "white",
                 domain = cropestab_map$perce_direct[cropestab_map$perce_direct>0], 
                 bins = c(0,25,50,75,100))

# pal <- colorBin(palette = "BuGn", domain = cropestab_map$direct, bins = c(0,1,10,25,50,100,Inf))

labels <- sprintf("Province: %s<br>Percent: %0.2f", str_to_title(cropestab_map$location), cropestab_map$perce_direct) %>%
  lapply(htmltools::HTML)

cropestab_map <- 
  cropestab_map |> 
  mutate(color = if_else(perce_direct==0, "white", pal(perce_direct)))

# Plot the map
leaflet(cropestab_map,
        width=700,
        height = 600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
  setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor = ~color,
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels
  ) %>%
  addLegend(pal = pal, values = ~perce_direct, 
            title = htmltools::HTML("Rice area planted with <br>direct seeding method (2022 DS)"), 
            opacity = 0.7,
              labFormat = function(type, cuts, p) {
    c("<25", "25-50", "50-75", "More than 75")  # matches the bins
  },
  na.label = "No adoption") |> 
      addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022</div>",
    position = "bottomleft"
  )
```
:::
:::

## Area % planted to Direct Seeding (RBFHS 2021 WS) {.tiny-text}

::: columns
::: {.column width="30%"}
::: flex-row
::: {.value-box .blue}
::: value-number
`r area_dssed_WS` %
:::

::: value-title
Area planted with Direct Seeding
:::
:::
:::
:::

::: {.column width="70%"}
```{r}
#|echo: false
#|fig-height: 12

cropestab_map2 <- 
province_map |> left_join(area_cropestab |> 
                            filter(year==2021,
                                   locType==2) |> 
                            mutate(location = clean_location(location)) |> 
                            select(location,totalarea:perce_direct)
                          )


bins2 <- quantile(cropestab_map2$perce_direct, probs = seq(0, 1, length.out = 5), na.rm = TRUE, names = FALSE)
pal2 <- colorBin(palette =c("#00F4FF",
                            "#00C5E9",
                            "#004CFF",
                            "#56A3F2"
                            ),
                 na.color = "white",
                 domain = cropestab_map2$perce_direct[cropestab_map2$perce_direct>0], 
                 bins = c(0,25,50,75,100))

labels <- sprintf("Province: %s<br>Percent: %0.2f", str_to_title(cropestab_map2$location), cropestab_map2$perce_direct) %>%
  lapply(htmltools::HTML)


cropestab_map2 <- 
  cropestab_map2 |> 
  mutate(color = if_else(perce_direct==0, "white", pal2(perce_direct)))

# Plot the map
leaflet(cropestab_map2,
        width = 700,
        height = 600) %>%
  addProviderTiles("CartoDB.VoyagerNoLabels",
                   options = providerTileOptions(minZoom =5)) %>%
   setView(lng = 121.7740+1, lat = 12.8797, zoom = 6) |> 
  addPolygons(
    fillColor = ~color,
    weight = 1,
    opacity = 1,
    color = "black",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 3, color = "#666", fillOpacity = 0.7),
    label = labels
  ) %>%
  addLegend(pal = pal, values = ~perce_direct, 
            title = htmltools::HTML("Rice area planted with <br>direct seeding method (2021 WS)"), 
            opacity = 0.7,
              labFormat = function(type, cuts, p) {
    c("<25", "25-50", "50-75", "More than 75")  # matches the bins
  },
  na.label = "No adoption") |> 
      addControl(
    html = "<div style='font-size: 12px;font-style: italic;'>Source:<br>RBFHS 2021-2022</div>",
    position = "bottomleft"
  )
```
:::
:::

# Rice Market

## Quarterly Farmgate Prices (Dry), 2010-2024 {.tiny-text}

```{r}
#| echo: false
library(openxlsx)
library(tidyverse)


riceprices <-  readRDS("riceprices.rds")


xriceprice <- 
  riceprices |> 
  filter(SERIES %in% c("New Series","2012-based","2018-based")) |> 
  filter(str_detect(COMMODITY,"WMR|RMR|Palay")) |> 
  filter(LOCATION=="PHILIPPINES") |> 
  mutate(date = ymd(paste(YEAR,PERIOD,"1",sep="-")),
         yr = year(date),
         mth = month(date),
         qtr = quarter(date),
         sem = semester(date))

xriceprice_QTR <- 
xriceprice |> 
  group_by(yr,qtr,TYPE,COMMODITY) |> 
  summarise_at(vars(PRICE),
               list(mean=mean,min=min,max=max)) |> 
  ungroup() |> 
  mutate(quarterdate = ymd(paste(yr,qtr*3,1,sep="-")),
         quartertext = paste(yr,"Quarter",qtr))
```

```{r}
#| label: farmgate
#| echo: false
#| 
plot_ly(x=~quarterdate,
        data = xriceprice_QTR |> 
          filter(str_detect(COMMODITY,"Palay \\(Other")),
        width=1000,
        height = 600) |> 
  add_lines(
    y = ~max,
    line = list(color = "transparent"),
    hoverinfo = 'none',
    showlegend = FALSE,
    name = "Highest Price") |>
  add_lines(
    y = ~min,
    line = list(color = "transparent"),
    fill = 'tonexty', fillcolor='rgba(0,100,80,0.2)',
    hoverinfo = 'none',
    showlegend = FALSE,
    name = "Lowest Price") |>

  add_trace(
    y = ~mean,
    type = 'scatter', mode = 'lines+markers',
    line = list(color = "darkgreen"),
    marker = list(color = "darkgreen"),
    text = ~paste("<b>",quartertext,"</b>",
                  "<br>Average: ₱",format(round(mean,digits = 2),big.mark=","),
                  "<br>Highest: ₱",round(max,digits=2),
                  "<br>Lowest: ₱", round(min,digits=2)),
    hoverinfo = 'text',
    name = "Average Farmgate Price (Dry)") |>
  layout(
    title = "Average, Highest and Lowest Farmgate  Prices (Dry), by Quarter",
    xaxis = list(title = "Year and Quarter",
                 rangeslider = list(visible = TRUE)),
    yaxis = list(title = "Pesos per kg (₱)"),
    hovermode = "x unified",
    # barmode = "stack",
    legend = list(x = 0.05, y = 1,
                  bgcolor = 'rgba(255, 255, 255, 0.4)')
    # hovermode = 'x unified'
  )|> 
    config(displayModeBar=FALSE)


```

## Quarterly Wholesale and Retail Rice Prices, 2010-2024 {.tiny-text}

```{r}
#| label: riceprice
#| echo: false

plot_ly(x=~quarterdate,
        width=1000,
        height = 600,
        data=xriceprice_QTR) |> 
  add_trace(data =xriceprice_QTR |> 
              filter(str_detect(TYPE,"Wholesale")) |> 
              filter(str_detect(COMMODITY,"WMR")) |> 
              filter(yr>=2012),
            x=~quarterdate,
            y=~mean,
            type = 'scatter',
            mode = 'lines',
            text = ~paste("<b>",quartertext,"</b>",
                          "<br> Wholesale - WMR: ₱", round(mean,digits=2)),
            hoverinfo='text',
            line = list(color ='darkgreen'),
            name = "Well Milled Rice (wholesale)") |> 
#Wholesale RTR
  add_trace(data =xriceprice_QTR |> 
              filter(str_detect(TYPE,"Retail")) |> 
              filter(str_detect(COMMODITY,"WMR")),
            x=~quarterdate,
        y=~mean,
        type = 'scatter',
        mode = 'lines',
        text = ~paste("Retail - WMR: ₱", round(mean,digits=2)),
        hoverinfo='text',
        line = list(color ='darkgreen',
                    dash = "dash"),
        name = "Well Milled Rice (retail)") |> 
  add_lines(
    data =xriceprice_QTR |> 
      filter(str_detect(TYPE,"Wholesale")) |> 
      filter(str_detect(COMMODITY,"RMR")) |> 
      filter(yr>=2012),
    y = ~mean,
    text = ~paste("Wholesale - RMR: ₱",round(mean,digits=2)),
    hoverinfo='text',
    line = list(color = "darkorange"),
    name = "Regular Milled Rice (wholesale)") |>
      add_lines(
    data =xriceprice_QTR |> 
      filter(str_detect(TYPE,"Retail")) |> 
      filter(str_detect(COMMODITY,"RMR")),
    y = ~mean,
    text = ~paste("Retail - RMR: ₱",round(mean,digits=2)),
    hoverinfo='text',
    line = list(color = "darkorange",
                dash = "dash"),
    name = "Regular Milled Rice (retail)") |>
  layout(
    title = "Average Wholesale and Retail Rice  Prices, by Quarter",
    xaxis = list(title = "Year and Quarter",
                 rangeslider = list(visible = TRUE)),
    yaxis = list(title = "Pesos per kg (₱)"),
    hovermode = "x unified",
    # barmode = "stack",
    legend = list(x = 0.05, y = 1,
                  bgcolor = 'rgba(255, 255, 255, 0.4)')
    # hovermode = 'x unified'
  )|> 
    config(displayModeBar=FALSE)


```

# Outlook for 2025 and beyond

## 2025 Areas with double dry

```{r}
#| label: doubleDry

bardata <- 
    nis_total |> 
    # filter(Category=="Total") |> 
    filter(comm=="Rice") |> 
    mutate(seas = recode_factor(seas,
                                "Dry1"="AUG to NOV<br> <i>(1st Dry)",
                                "Dry2"="DEC to MAR<br> <i> (2nd Dry)",
                                "Ddry"="<b>Double Dry",
                                "Wet"="APR to JUL<br> <i> (Regular Wet)",
                                .ordered = T)
           )

seas_order <- levels(bardata$seas)
#split

bardata_ddry <- 
  bardata |> filter(seas == "<b>Double Dry")

bardata_xddry <- 
  bardata |> filter(seas!= "<b>Double Dry")


plot_ly(width = 1000, height = 600) %>%
  # Other seasons: opacity 0.6
  add_trace(
    data = bardata_xddry,
    x = ~val,
    y = ~rev(seas),
    type = 'bar',
    orientation = 'h',
    color = ~Category,
    colors = c("#A0522D", "#B68B00", "darkgreen"),
    opacity = 0.4,
    text = ~paste(format(round(val), big.mark = ","), "HA"),
    textposition = 'inside',
    insidetextanchor = 'middle',
    hoverinfo = 'none',
    textfont = list(color = 'white', size = 15),
    showlegend = TRUE
  ) %>%
  # Ddry: full opacity
  add_trace(
    data = bardata_ddry,
    x = ~val,
    y = ~rev(seas),
    type = 'bar',
    orientation = 'h',
    color = ~Category,
    colors = c("#A0522D", "#B68B00", "darkgreen"),
    opacity = 1,
    text = ~paste(format(round(val), big.mark = ","), "HA"),
    textposition = 'inside',
    insidetextanchor = 'middle',
    hoverinfo = 'none',
    textfont = list(color = 'white', size = 15),
    showlegend = FALSE  # prevent double legend for same categories
  ) %>%
  layout(
    title = 'Total NIS and CIS program area for rice, by season, CY 2025',
    yaxis = list(
      title = "",
      categoryorder = "array",
      categoryarray = rev(seas_order)
    ),
    xaxis = list(
      title = "Program area (HA)"
    )
  ) %>%
  config(displayModeBar = 0)



# plot_ly(
#   data=nis_total |> 
#     # filter(Category=="Total") |> 
#     filter(comm=="Rice") |> 
#     mutate(seas = recode_factor(seas,
#                                 "Dry1"="AUG to NOV<br> <i>(1st Dry)",
#                                 "Dry2"="DEC to MAR<br> <i> (2nd Dry)",
#                                 "Ddry"="<b>Double Dry",
#                                 "Wet"="APR to JUL<br> <i> (Regular Wet)",
#                                 .ordered = T)
#            ),
#   width=1000,
#   height=600,
#   x=~val,
#   y=~rev(seas),
#   color = ~Category, #"#F0C419"
#   colors =  c("#A0522D", "#B68B00","darkgreen"),
#   # colors = c("#6E3B1A","#B68B00","#2E6B13"),
# 
#   # marker = list(color = "darkgreen"),
#   type = 'bar',
#   orientation = 'h',
#   text = ~paste(format(round(val), big.mark=","), "HA"),                      # Add value as text
#   textposition = 'inside',         # Position text inside the bar
#   insidetextanchor = 'middle',  
#   hoverinfo = 'none',
#   # Align text at the start of the bar
#   textfont = list(color = 'white', size = 15)) |> # Optional: style the text) |> 
#   layout(title = 'Total NIS and CIS program area for rice, by season, CY 2025',
#          yaxis = list(
#            title = "",
#            categoryorder = "array",
#            categoryarray = unique(nis_total$seas)
#          ),
#          xaxis = list(
#            title = "Program area (HA)"
#          )
#   ) |> 
#   config(displayModeBar=0)
#   


```

<!-- ## SUA projections {.tiny-text} -->

<!-- ::: tiny-text -->

<!-- ```{r} -->

<!-- #| label: sua projected -->

<!-- #| echo: false -->

<!-- ensemb_form <- readRDS("esembform.rds") -->

<!-- sua_ensemb<-  -->

<!-- ensemb_form[[".supply"]] %>% -->

<!--   filter(!(variables %in% c("Base Palay Production", -->

<!--                             "Total Calamity loss"))) %>%  -->

<!--   rbind(ensemb_form[[".imports"]]) %>%  -->

<!--   mutate(group ="SUPPLY") %>%  -->

<!--   rbind(ensemb_form[[".demand"]] %>%  -->

<!--           mutate(group ="DEMAND")) %>%  -->

<!--   rbind(ensemb_form[[".ssr"]] %>%  -->

<!--           mutate(group = " ")) |>  -->

<!--   datatable(rownames = FALSE, -->

<!--             options = list(pageLength = 20, -->

<!--                            ordering=FALSE, -->

<!--                            dom='t',  -->

<!--                            # scrollX = TRUE, -->

<!--                            # buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), -->

<!--                            # rowGroup = list(dataSrc = 7), -->

<!--                            columnDefs = list( -->

<!--                              list(visible = FALSE, targets=8) -->

<!--                              # list(targets=2:9, class="dt-right") -->

<!--                            ) -->

<!--             ), -->

<!--             class = "compact nowrap stripe hover" -->

<!--   ) %>%  -->

<!--   formatStyle(3:9, -->

<!--               `text-align` = "right") %>%  -->

<!--   formatCurrency(3:9, -->

<!--                  digits = 0, -->

<!--                  currency ="", -->

<!--                  interval=3, -->

<!--                  rows = 9:18, -->

<!--                  mark=",") -->

<!-- sua_ensemb -->

<!-- ``` -->

<!-- -   SUA projections 2025 - 2030 based on DAC Ensemble models -->

<!-- ::: -->

## SUA Target 2025-2030 {.tiny-text}

::: tiny-text
```{r}
#| label: sua_target
#| echo: false


sua_new5_form <-  readRDS("sua_new5_form.rds")
sua_target <- 
  sua_new5_form[[".supply"]] %>%
  filter(!(variables %in% c("Base Palay Production",
                            "Total Calamity loss"))) %>% 
  rbind(sua_new5_form[[".imports"]]) %>% 
  mutate(group ="SUPPLY") %>% 
  rbind(sua_new5_form[[".demand"]] %>% 
          mutate(group ="DEMAND")) %>% 
  rbind(sua_new5_form[[".ssr"]] %>% 
          mutate(group = " ")) |> 
  select(-`2024`) |> 
  datatable(rownames = FALSE,
            # extensions = "Buttons",
            options = list(pageLength = 30,
                           ordering=FALSE,
                           dom='t', 
                           # buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           # rowGroup = list(dataSrc = 7),
                           columnDefs = list(
                             list(visible = FALSE, targets=8:9)
                             # list(targets=2:9, class="dt-right")
                           )
            ),
            class = "compact nowrap stripe hover"
  ) %>% 
  formatStyle(3:9,
              `text-align` = "right") %>% 
  formatCurrency(3:9,
                 digits = 0,
                 currency ="",
                 interval=3,
                 rows = 9:18,
                 mark=",")

sua_target
```

-   SUA targets: Reach 100% self sufficiency by 2030 at *63% MR* and *122.7 kg/year per capita consumption*.
:::

## Reaching Sufficiency through demand management {.tiny-text}

```{r}
#| echo: false

suff_percap <-  read_csv("diffpercap.csv") |> 
  pivot_longer(-c(mr,percap),
               names_to = c("var","year"),
               names_sep = '_',
               values_to = "value") |> 
  mutate(Yr = factor(year,levels = c(2024,2025,2026,2027,2028,2029,2030)
  )) |> 
  pivot_wider(names_from=var,
              values_from=value)



plot_ly(data = suff_percap |> 
          filter(mr==0.63),
        x=~Yr,
        height=600,
        width=1000) |> 
  add_lines(y=~`Sufficiency Ratio`,
            color = ~factor(percap)) |> 
  layout(
    title =  " Rice Self Sufficiency levels at different Per capita consumption",
    yaxis = list(title = "Rice Self Sufficiency Ratios (%)"),
    hovermode = 'x unified',
    xaxis = list(title = "Year"),
    legend = list(x = 0.05, y = 0.95,
                  title = list(text = "<b>Per Capita Consumption (kg/yr)</b>"),
                  bgcolor = 'rgba(255, 255, 255, 0.4)',
                      orientation = "v",                     # vertical orientation (default)
                  tracegroupgap = 2,                     # optional spacing between groups
                  xanchor = "left",
                  x = 1.02,
                  y = 1,
                  itemsizing = "constant",
                  traceorder = "normal"),
    margin = list(r = 80)
    ) |> 
    config(displayModeBar=FALSE)

```

## Needed government investment to provide 20 P/kg Rice in the market {.extra-tiny-text}

::: tiny-text
```{r}

price_scenarios <- read.xlsx("peso20budget.xlsx",
                             sheet=1) |> 
  mutate_if(is.numeric,
            round,digits=2) |> 
  mutate_if(is.numeric,
            format,nsmall=2)

budget_scenarios <- read.xlsx("peso20budget.xlsx",
                              sheet=2) |>
  mutate_if(is.numeric,
            round,digits=2) |> 
  mutate_if(is.numeric,
            format,nsmall=2)


price_scenario_labels <- data.frame(Item = "Farmgate and Wholesale prices",
                                    A=NA,
                                    B=NA,
                                    C=NA,
                                    D=NA)

budget_scenario_labels <- data.frame(Item = "Required budget to subsidize the following quantity scenarios",
                                    A=NA,
                                    B=NA,
                                    C=NA,
                                    D=NA)

merged_x <- data.frame(
  Item = 'Required budget to subsidize the following quantity scenarios:',
  A = 'In Billion Pesos',
  B = '',
  C = '',
  D = ''
)

combined_table <- 
  rbind(
  price_scenarios,
  "",
  merged_x,
  budget_scenarios
)

  
datatable(
  combined_table,
  rownames=FALSE,
  container = htmltools::withTags(table(
    class = 'display',
    thead(
      tr(
        th(colspan = 1, "Farmgate and Wholesale prices"),
        th(colspan = 4, "Price per kg")
      )
    )
  )),
    options = list(
      paging = FALSE,
      dom='t',
          headerCallback = JS(
      "function(thead, data, start, end, display) {",
      "  // Only center the second merged header cell",
      "  $(thead).find('th').eq(1).css({'text-align': 'center'});",
      "}"
    ),
                rowCallback = JS(
              "function(row, data, index) {",
              "  if(data[0] === 'Required budget to subsidize the following quantity scenarios:') {",
              "    // Style A cell",
              "    $('td', row).eq(0).css({'font-weight':'bold'});",
              "    // Merge A to D",
              "    $('td', row).eq(1).attr('colspan', 4);",
              "    $('td', row).eq(1).css({'text-align':'center', 'font-weight':'bold'});",
              "    // Hide the other cells after merge",
              "    $('td', row).eq(2).hide();",
              "    $('td', row).eq(3).hide();",
              "    $('td', row).eq(4).hide();",
              "  } else {",
              "    // Indent values in column A (except 'Group A')",
              "    $('td', row).eq(0).css({'padding-left': '20px'});",
              "  }",
              "}"
            )
    )
)
```
:::

::: tiny-text
-   Per Capita consumption at 122.7kg/year
-   2025 population at 115 Million (Source: PSA)
-   Poverty incidence at 15.5 percent (Source: PSA)
-   Subsistence incidence at 4.3 percent (Source: PSA)
:::

## 

::: thank-you
::: {style="font-size: 2.5em; font-weight: bold; text-align: center;"}
Thank You!!
:::

::: {style="margin-top: 1em; font-size: 1.2em; text-align: center;"}
Questions or feedback?<br> <strong>philrice.dataanalytics\@gmail.com</strong>
:::
:::
